Modelling the emergence of whisker barrels

### 2020-06-16 paper/briefpaper/brief.tex
Localization mostly replaced with selectivity

A cortical boundary enclosing 41 \cmnt{macrovibrissae} barrels was traced from
a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
gradients aligned with the anterior-posterior and medial-lateral axes. From
random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
  $c(\mb{x},0)=0$}, a clear Voronoi-like tesselation
emerged (Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta
metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
pattern emerged within \cmnt{$\approx$~20000} iterations
(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
between real and simulated barrels revealed a strong correspondence (see
Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
  \emph{the connection density of the TC type $i$ which makes the densest
    connections} to \emph{the sum of all connection densities} gives a measure
  of the localization of the connection density. We computed this measure
  across the entire barrel domain and named it $\omega$ (see
  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
  20000 simulation steps and converges on a stable
  value. Fig.\,\ref{fig:main}E shows the localization measure plotted as a
  luminance map, giving a more graded impression of the structure which
  develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\cmnt{\textbf{A} Exploring the parameter space of the model given
        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
        barrel area difference and $\omega$ (bottom row) is a measure of
        localization (separation) of the connection densities for the 41
        thalamocortical axon types. For each row, lighter (orange/yellow)
        colours imply a better model. The parameter space explored is three
        dimensional with the parameter ratio $\alpha/\beta$ varied along the
        columns of graphs, the competition parameter $\epsilon$ varied on the
        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
        asterisk (*) marks the parameter set displayed in
        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\cmnt{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
  iterations plotted as colour maps for $216$ parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows
  results for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$ varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
  are yellow. Lower values of the difference between real and simulated barrel
  area indicates good pattern matching. Higher values of $\omega$ indicate
  better localization and are coloured yellow. White squares indicate that the
  simulation was not stable for that parameter set.}

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the maximally
  localized solution, which for $D$ should be small. Thus, there is a
  trade-off between good localization and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
  of varying $D$ in either direction from its optimum. ii)~is a less
  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

\cmnt{We performed a sensitivity analysis to determine if the quality of the
  pattern was affected by perturbations to any of the following inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
%
  In the following analysis, we always evaluated the three pattern metrics,
  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
  $\omega$ (which measures localization of connections, $c_i$) after running
  the model for 30000 simulation steps.}

@@

\cmnt{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
  first column shows the results for the three metrics where the interaction
  parameters $\gamma_{i,j}$ are replaced by $\gamma_{i,j}'$ according to
  $\gamma_{i,j}' = \gamma_{i,j} + U(0,4\nu_\gamma)$, where $\nu_\gamma$ is the
  chosen magnitude of the noise (the proportion of the full range of
  $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
  and the localization of the fields are not strongly affected by interaction
  parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}

>>>>

A cortical boundary enclosing 41 \cmnt{macrovibrissae} barrels was traced from
a cytochrome oxidase stain from \cite{zheng_signal_2001} \cmnt{(using the
  authors' original data)}, and Eqs.~\ref{eq:dc}--\ref{eq:comp} were solved
for $N=41$ projections on the resulting domain, using $M=2$ linear signalling
gradients aligned with the anterior-posterior and medial-lateral axes. From
random initial conditions for $a(\mb{x},0)$, \cmnt{with each hexagonal element
  of $a_i$ set from a uniform distribution according to $0.2+{U}(0,0.2)$, and
  $c(\mb{x},0)=0$}, a clear Voronoi-like tesselation
emerged (Fig.\,\ref{fig:main}C; see Movie S1). A reduction in the Honda delta
metric (see \citealp{senft_mouse_1991}) confirmed that a `good' Voronoi
pattern emerged within \cmnt{$\approx$~20000} iterations
(Fig.\,\ref{fig:main}D).  A measure of the difference in \cmnt{barrel area}
between real and simulated barrels revealed a strong correspondence (see
Fig.\,\ref{fig:main}D, squares). \cmnt{In a chosen region, the ratio of
  \emph{the connection density of the TC type $i$ which makes the densest
    connections} to \emph{the sum of all connection densities} gives a measure
  of the selectivity of the connection density, $\mu(\mb{x},t)$. We summed this measure
  across the entire barrel domain and named it $\omega(t)$ (see
  Fig.\,\ref{fig:main}D, grey hexagons). $\omega$ increases during the first
  20000 simulation steps and converges on a stable
  value. Fig.\,\ref{fig:main}E shows the selectivity measure plotted as a
  luminance map, giving a more graded impression of the structure which
  develops in the system.}

% Full page width
\begin{figure}
  \begin{fullwidth}
    \includegraphics[width=\linewidth]{./Fig2.png}
    \caption{\cmnt{\textbf{A} Exploring the parameter space of the model given
        in Eqs.\,1--3. Graphs show the three pattern metrics for $t=30000$
        steps as colourmaps: Honda $\delta$ (top row) shows conformance to a
        Voronoi pattern; $\eta$ (middle row) gives the simulated/experimental
        barrel area difference and $\omega$ (bottom row) is a measure of
        selectivity of the connection densities for the 41
        thalamocortical axon types. For each row, lighter (orange/yellow)
        colours imply a better model. The parameter space explored is three
        dimensional with the parameter ratio $\alpha/\beta$ varied along the
        columns of graphs, the competition parameter $\epsilon$ varied on the
        $y$-axes and the diffusion constant, $D$ varied along the $x$-axes. An
        asterisk (*) marks the parameter set displayed in
        Fig.\,\ref{fig:main}. Boxes i), ii)~and iii)~mark parameter sets for which
        corresponding patterns are shown in \textbf{B} (for $t=30000$ steps).}}
    \label{fig:paramsweep}
  \end{fullwidth}
\end{figure}

\cmnt{To determine the robustness of the model to its parameters, we
  investigated its behaviour with respect to changes in the parameters $D$,
  $\epsilon$ and the ratio of $\alpha$ to $\beta$. Fig.\,\ref{fig:paramsweep}A
  shows the value of the metrics $\delta$, $\eta$ and $\omega$ after $30000$
  iterations plotted as colour maps for $216$ parameter sets. Each row of maps
  shows the results for a separate metric. Each column of maps shows
  results for a different value of the ratio $\alpha / \beta$. The diffusion
  parameter, $D$ varies with the $x$ axes and the strength of competition,
  $\epsilon$ varies with the $y$ axes. The colour maps are arranged so that
  `yellow is good'. Lower values of $\delta$ indicate good Voroni patterns and
  are yellow. Lower values of the difference between real and simulated barrel
  area indicates good pattern matching. Higher values of $\omega$ indicate
  greater selectivity and are coloured yellow. White squares indicate that the
  simulation was not stable for that parameter set.}

\cmnt{The maps indicate that the pattern is not strongly influenced by the
  choice of $\alpha/\beta$, which was varied by 5 orders of magnitude. The
  most Voronoi-like patterns were achieved for parameter sets for which $D$
  and $\epsilon$ were high. The barrel area metric, $\eta$ indicates a strong
  preference for $D\approx0.5$, although this does not give the
  solution of maximum selectivity, which for $D$ should be small. Thus, there is a
  trade-off between good selection and good Voroni pattern
  conformation. The
  parameter set indicated by the asterisk in
  Fig.\,\ref{fig:paramsweep}A is presented in Fig.\,\ref{fig:main} because it
  provides the best barrel area correspondence and produces a good Voronoi
  pattern. In Fig.\,\ref{fig:paramsweep}B, we present some patterns
  representing alternative parameter choices. Fig.\,\ref{fig:paramsweep}B i)
  shows the simulation result with low diffusion constant, $D$. Reducing the
  diffusion constant is equivalent to increasing the size of the domain $S$
  and so this simulation can be thought to represent an animal with a large
  cortical area but with unchanged cortical complexity. In this simulation, a
  large proportion of the domain is occupied by the projections with
  intermediate interaction parameters; those with strong interaction
  parameters are bunched up towards the edge of the domain. A barrel pattern
  fails to form. Fig.\,\ref{fig:paramsweep}B ii)~and iii)~indicate the effect
  of varying $D$ in either direction from its optimum. ii)~is a less
  exaggerated version of the pattern in i), whereas iii)~shows an expansion of
  the outer barrels and a corresponding compression of the barrels within.}

@@

\cmnt{We performed a sensitivity analysis to determine if the quality of the
  pattern was affected by perturbations to any of the following inputs:
%
  i)~the magnitude and offset of the noise applied to $a_i$ at $t=0$;
%
  ii)~random noise applied to the interaction parameters, $\gamma_{i,j}$;
%
  iii)~random noise (at various length scales) applied to the guidance fields;
%
  and
%
  iv)~The angle and magnitude of one of the guidance fields, ($\rho_1$;
  Fig.\,\ref{fig:main}A).
%
  In the following analysis, we always evaluated the three pattern metrics,
  $\delta$ (the Honda Voronoi metric), $\eta$ (the pattern-match metric) and
  $\omega$ (which measures selectivity of connections, $c_i$) after running
  the model for 30000 simulation steps.}

@@

\cmnt{Fig\,\ref{fig:sens}A shows the results for perturbations ii) to iv). The
  first column shows the results for the three metrics where the interaction
  parameters $\gamma_{i,j}$ are replaced by $\gamma_{i,j}'$ according to
  $\gamma_{i,j}' = \gamma_{i,j} + U(0,4\nu_\gamma)$, where $\nu_\gamma$ is the
  chosen magnitude of the noise (the proportion of the full range of
  $\gamma_{i,j}$, which is 4). The Voronoi nature of the pattern ($\delta$)
  and the selectivity of the fields are not strongly affected by interaction
  parameter noise, but the pattern metric, $\eta$ shows a... ADD SUB GRAPH.}
